$\pi$SOD-M: A Methodology for Building Service-Oriented Applications in the Presence of Non-Functional Properties

### 2014-02-12 JournalToSubmit2013/metamodels.tex
section 3

Extending from the highest level of abstraction of the MDA, $\pi$-SODM provides  a conceptual structure to: first, capture the system requirements and specification in high-level abstraction models (computation independent models, CIMs); next,  starting from such models build platform independent models (PIMs) specifying the system details; next transform such models into platform specific models (PSMs) that bundles the specification of the system with the details of the targeted platform; and finally, serialize such model into code that implements the system.

\subsubsection{Computation Independent Models}
$\Pi$-SODM uses two models at the CIM level: The E3value~\cite{e3value} model and BPMN~\cite{BPMN}.
The former  identifies the transference of value information between components of the system.
The BPMN model establishes which are the actors and main tasks of the application.

The e3value defines \textit{dependency paths}, showing the value exchanges, which are triggered by the occurrence of an end-consumer need (in our case, the need of a risk assessment). 
A dependency path has a direction and consists of a sequence of linked dependency nodes.
A dependency path starts with a \textit{start stimulus} node and ends with an \textit{end stimulus} node\footnote{See Legend on Figure~\ref{fig:E3valuemodel}}. 
Dependency paths may also contain \textsl{OR} and \textsl{AND} elements (both for initiate and join alternative and parallel paths).

\subsubsection{Platform Independent Models}
 The \textit{$\pi$-UseCase} meta-model extends the UML Use Case meta-model for describing functional and non-functional requirements. Non-functional requirements are defined as \textit{constraints} over actions and data.

@@

\begin{example}[To Publish Music]\label{ex:toPublicMusic}
To illustrate the use of the $\pi$-Service Composition Meta-model, we define a model for the ``To Publish Music'' scenario (Figure \ref{fig:servicecompositionmodel}). 
In this model, there are three external business collaborators ({\em Spotify, Twitter} and {\em Facebook}).
% \footnote{We use {\em italics} to refer to concrete values of the classes of a model that are derived from the classes of a meta-model.}). 
The model also shows the business process of the application that consists of three service activities: {\em Listen Music}, {\em Public Music} and {\em Confirmation}. 
Note that  the activity {\em Publish Music} calls the actions of two service collaborators namely {\em Facebook} and {\em Twitter}.
Both {\em Facebook} and {\em Twitter} services require authentication protocols in order to execute methods that will read and update the user space. 
%A call to such services must be part of the authentication protocol required by these services.
In the example we  associate two authentication policies, one for the open authentication protocol, represented by the class {\sf\small OAuthPolicy} at {\em Twitter}, that will be associated to the activity  {\sf\small UpdateTwitter} (see Figure \ref{fig:servicecompositionmodel}). 
In the same way, the {\em Facebook} class {\sf\small HTTPAuthPolicy}, for the http authentication protocol will be associated to the activity {\sf\small UpdateFacebook}.

>>>>

%Extending from the highest level of abstraction of the MDA, $\pi$-SODM provides  a conceptual structure to: first, capture the system requirements and specification in high-level abstraction models (computation independent models, CIMs); next,  starting from such models build platform independent models (PIMs) specifying the system details; next transform such models into platform specific models (PSMs) that bundles the specification of the system with the details of the targeted platform; and finally, serialize such model into code that implements the system.

\subsection{Computation Independent Models}
$\Pi$-SODM uses two models at the CIM level: The e$^3$value~\cite{e3value} model and BPMN~\cite{BPMN}.
The former identifies the value/information exchange between components of the system.
The BPMN model establishes which are the actors and main tasks of the application.

The e$^3$value model defines \textit{dependency paths}, showing the value exchanges, which are triggered by the occurrence of an end-consumer need (in our case, the need of a risk assessment). 
A dependency path has a direction and consists of a sequence of linked dependency nodes.
A dependency path starts with a \textit{start stimulus} node and ends with an \textit{end stimulus} node\footnote{See Legend on Figure~\ref{fig:E3valuemodel}}. 
Dependency paths may also contain \textsl{OR} and \textsl{AND} elements (both for initiate and join alternative and parallel paths).

The next example ilustrates the use of the models.

\begin{example}[To Publish Music]\label{ex:toPublicMusic}
To illustrate our proposal, we build the models for a ``To Publish Music'' problem. 
This problem involves three actors ({\em Spotify, Twitter} and {\em Facebook}).
We model a business process that consists of three service activities: {\em Listen Music}, {\em Publish Music} and {\em Confirmation}. 
The {\em Publish Music} activity calls the {\em Facebook} and {\em Twitter} services.
Both {\em Facebook} and {\em Twitter} require authentication. 
Two authentication policies are required, one for {\em Twitter} and another for {\em Facebook}.

THe e$^3$value and BPMN models for this applications are shown in Figure~\ref{fig:CIM:tpm}.
\hfill\openbox
\end{example}

\begin{figure}
\framebox{\color{red}\Huge Valeria, please add the figures?}
\caption{\label{fig:CIM:tpm} CIM models for ``To Publish Music''.}
\end{figure}

In the next sections, we present each meta-model of $\pi$SOD-M, the models for the ``To Publish Music'' example, as well as the transformations between models.
Notice that \textit{ad-hoc} transformations are used to obtain PIM models from the CIM models proposed here.

\subsection{Platform Independent Models}
 The \textit{$\pi$-UseCase} meta-model extends the UML Use Case meta-model for describing functional and non-functional requirements. Non-functional requirements are defined as \textit{constraints} over actions and data.

@@

\begin{example}[To Publish Music]\label{ex:toPublicMusic}
To illustrate the use of the $\pi$-Service Composition Meta-model, we define a model for the ``To Publish Music'' scenario (Figure \ref{fig:servicecompositionmodel}). 
In this model, there are three external business collaborators ({\em Spotify, Twitter} and {\em Facebook}).
% \footnote{We use {\em italics} to refer to concrete values of the classes of a model that are derived from the classes of a meta-model.}). 
The model also shows the business process of the application that consists of three service activities: {\em Listen Music}, {\em Publish Music} and {\em Confirmation}. 
Note that  the activity {\em Publish Music} calls the actions of two service collaborators namely {\em Facebook} and {\em Twitter}.
Both {\em Facebook} and {\em Twitter} services require authentication protocols in order to execute methods that will read and update the user space. 
%A call to such services must be part of the authentication protocol required by these services.
In the example we  associate two authentication policies, one for the open authentication protocol, represented by the class {\sf\small OAuthPolicy} at {\em Twitter}, that will be associated to the activity  {\sf\small UpdateTwitter} (see Figure \ref{fig:servicecompositionmodel}). 
In the same way, the {\em Facebook} class {\sf\small HTTPAuthPolicy}, for the http authentication protocol will be associated to the activity {\sf\small UpdateFacebook}.

### 2014-02-12 JournalToSubmit2013/modelingFlyingPig.tex
section 3

\begin{figure}[t]
\centering
%\includegraphics[width=1.0\textwidth]{figs/ServiceCompositionGeneralCut.png}
\fbox{\LARGE $\pi$-PEWS Model.}

{\color{red} \raggedright
$\star$ Placido: Please, place the model here.
}
\caption{$\pi$-PEWS model for \FlyingPig.\label{fig:PiPEWSModel}}
\end{figure}

>>>>

%\begin{figure}[t]
%\centering
\begin{lstlisting}[label=list:flyingPigPiPEWS,caption=pi-PEWS
Specification: FlyingPig. ]!\label{fig:PiPEWSModel}!

//Namespaces specify service URI
namespace orca = www.orca.mx/service.wsdl

//Operations
alias createCompanySpace = portType/createCompanySpace in orca
alias generateQuestionnaire = portType/generateQuestionnaire in orca
alias analizeAnswer = portType/analizeAnswer in orca
alias storeData = portType/storeData in orca
alias createReport = portType/createReport in orca
alias callForAmendments = portType/callForAmendments in orca

//Services
service receiveRequest(R, Id) = createCompanySpace(R, Id)

service generateNewInterface(Id, NULL) = ...

service createQuestionnaire(Id, Q) = generateQuestionnaire(Id, Q)

service notifyResponsable((Id, Q); NULL) = ...

service receiveAnswers((Id, T); P) =
			analizeAnswer((Id, T), NULL) 
			. storeData((Id, T), NULL)
			. (
			  (createReport(Id, P) . return(P))
			  + (callForAmendments(Id,T) . return(NULL))
			  )

//Path

receiveRequest(R, Id)
||
( generateNewInterface(Id, NULL)
  . createQuestionnaire(Id, Q)
  .  notifyResponsable((Id, Q); NULL
 )
||
 ( receiveAnswers((Id, T); P)
 . [P != NULL] STOP
 )*

//Policies

\end{lstlisting}
%\fbox{\LARGE $\pi$-PEWS Model.}
%
%{\color{red} \raggedright
%$\star$ Placido: Please, place the model here.
%}
%\caption{$\pi$-PEWS model for \FlyingPig.\label{fig:PiPEWSModel}}
%\end{figure}

### 2014-02-12 JournalToSubmit2013/modelingWithPISODM.tex
section 3

In this section we present $\pi$SOD-M, an MDD based methodology for building service compositions with non-functional requirements. $\pi$SOD-M provides concepts for  modelling  non-functional requirements at the early stages of the development.
%The method uses the concept of \textit{Policy}~\cite{Espinosa-Oviedo2011a} to represent non-functional requirements.

@@

%$\pi$-SODM meta-models represent both the functional aspects of the application as well as its non-functional constraints.
As shown in Figure \ref{fig:piSOD-M}, according to MDA guidelines, $\pi$-SODM proposes meta-models  organized into three levels: 
\begin{itemize}
\item CIM (\textit{Computational Independent Models}). This level focusses on the
environment of the system, as well as on its business and requirement specifications.
This level represents the software system at its highest level of abstraction.
At this stage of the development, the structure and system processing details are still unknown or undetermined.  
$\pi$-SODM uses the \textit{E3value}~\cite{e3value} and \textit{BPMN}~\cite{BPMN} meta-models for this purpose. 
%Referring to the "To Publish Music" application,  the designer starts defining an E3value model \footnote{The E3 value model is a business model that represents a business case %graphically as a set of value exchanges ($\nabla$$\triangle$) and value activities (rounded boxes) performed by business actors (squared boxes) 
%and allows  to understand the environment in which the services' composition will be placed \cite{e3value}.}  at the CIM level and then the corresponding models of the PIM are generated leading to a services' composition model (SCM).

>>>>

In this section we present $\pi$SOD-M, an MDD-based methodology for building service compositions with non-functional requirements. 
$\pi$SOD-M extends SOD-M with concepts for modelling non-functional requirements.
This is done at all levels of MDA, starting from the early stages of the development.
%The method uses the concept of \textit{Policy}~\cite{Espinosa-Oviedo2011a} to represent non-functional requirements.
%$\pi$-SOD-M meta-models represent both the functional aspects of the application as well as its non-functional constraints.

@@

$\pi$-SODM proposes meta-models organized into three levels (Figure~\ref{fig:piSOD-M}): 
\begin{itemize}
\item CIM (\textit{Computational Independent Models}). This level focusses on the
environment of the system, as well as on its business and requirement specifications.
This level represents the software system at its highest level of abstraction.
At this stage of the development, the structure and system processing details are still unknown or undetermined.  
$\pi$-SODM uses the \textit{e$^3$value}~\cite{e3value} and \textit{BPMN}~\cite{BPMN} meta-models for this purpose. 
%Referring to the "To Publish Music" application,  the designer starts defining an E3value model \footnote{The E3 value model is a business model that represents a business case %graphically as a set of value exchanges ($\nabla$$\triangle$) and value activities (rounded boxes) performed by business actors (squared boxes) 
%and allows  to understand the environment in which the services' composition will be placed \cite{e3value}.}  at the CIM level and then the corresponding models of the PIM are generated leading to a services' composition model (SCM).
